---
title: "Grading Rubric for Academic Theses"
author: "Prof. Dr. Claudius Graebner"
date: today
date-modified: "2026-02-18"
modified-title: "Last modified"
date-format: "MMMM DD, YYYY"
lang: en
format:
  html:
    toc: true
    toc-depth: 2
    theme: cosmo
    embed-resources: true
  pdf:
    documentclass: scrartcl
    papersize: a4
    fontsize: 11pt
    geometry: "margin=2.5cm"
    toc: false
  docx:
    toc: true
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
library(flextable)
library(officer)

# Color palette
clr_header <- "#1B3A5C"
clr_header_text <- "#FFFFFF"
clr_light <- "#F0F4F8"
clr_accent <- "#2E6EA6"

# Consistent table theme
theme_rubric <- function(ft) {
  ft |>
    fontsize(size = 10, part = "all") |>
    font(fontname = "Arial", part = "all") |>
    bg(bg = clr_header, part = "header") |>
    color(color = clr_header_text, part = "header") |>
    bold(part = "header") |>
    align(align = "left", part = "all") |>
    border_remove() |>
    hline_top(border = fp_border(color = clr_header, width = 2), part = "header") |>
    hline_bottom(border = fp_border(color = clr_header, width = 2), part = "header") |>
    hline_bottom(border = fp_border(color = clr_header, width = 1.5), part = "body") |>
    padding(padding.top = 4, padding.bottom = 4, part = "all") |>
    autofit()
}
```

## Overview

Your thesis is assessed along seven dimensions, each weighted according to its
importance. Every dimension is scored on a scale from 0 to 3 points.
The overall grade results from the weighted sum of all dimension scores.

```{r}
#| label: tbl-overview
#| tbl-cap: "Assessment dimensions and weights"

overview <- data.frame(
  No = 1:7,
  Dimension = c(
    "Argumentation & Logic",
    "Research Design & Methods",
    "Literature & Theory",
    "Results & Analysis",
    "Critical Reflection",
    "Academic Craft",
    "Independence & Difficulty"
  ),
  Code = c("ARG", "MET", "LIT", "RES", "REF", "CRF", "IND"),
  Weight = c("20%", "15%", "15%", "20%", "10%", "15%", "5%")
)

flextable(overview) |>
  theme_rubric() |>
  align(j = c("No", "Code", "Weight"), align = "center", part = "all") |>
  width(j = "Dimension", width = 3.2) |>
  width(j = "Weight", width = 0.8)
```

## Scoring Scale

```{r}
#| label: tbl-scale
#| tbl-cap: "Scoring scale"

scale_df <- data.frame(
  Score = c("3", "2", "1", "0"),
  Rating = c("Excellent", "Good", "Adequate", "Insufficient"),
  Description = c(
    "Clearly exceeds expectations; outstanding quality",
    "Fully meets expectations with solid quality throughout",
    "Meets basic requirements but with noticeable weaknesses",
    "Does not meet requirements; substantial deficiencies"
  )
)

flextable(scale_df) |>
  theme_rubric() |>
  align(j = "Score", align = "center", part = "all") |>
  width(j = "Description", width = 4.0)
```

For dimensions where nothing particularly noteworthy stands out --- neither positively
nor negatively --- a default score of 1.5 points is assigned. The grade therefore
reflects deviations from this average, both strengths and weaknesses.

## Dimensions in Detail

### 1. Argumentation & Logic (ARG, 20%)

This dimension assesses the quality and persuasiveness of your reasoning.

```{r}
#| label: tbl-arg

arg_df <- data.frame(
  Criterion = c(
    "Clear thesis / research question",
    "Logical coherence of argument",
    "Claims supported by evidence",
    "Conclusion follows from analysis"
  ),
  Description = c(
    "Is there a precise, well-delimited research question or thesis statement?",
    "Do the arguments follow logically and without contradiction?",
    "Are claims backed up with sources, data, or examples?",
    "Does the conclusion follow naturally from the preceding analysis?"
  )
)

flextable(arg_df) |>
  theme_rubric() |>
  width(j = "Criterion", width = 2.8) |>
  width(j = "Description", width = 3.8)
```

### 2. Research Design & Methods (MET, 15%)

This dimension assesses whether the methodology fits the research question and is
presented transparently.

```{r}
#| label: tbl-met

met_df <- data.frame(
  Criterion = c(
    "Method appropriate to question",
    "Method clearly explained",
    "Limitations acknowledged",
    "Reproducibility / transparency"
  ),
  Description = c(
    "Does the chosen method suit the research question and the material?",
    "Is the procedure described clearly enough for others to follow?",
    "Are methodological limitations identified and discussed?",
    "Could other researchers reproduce the analysis with the information provided?"
  )
)

flextable(met_df) |>
  theme_rubric() |>
  width(j = "Criterion", width = 2.8) |>
  width(j = "Description", width = 3.8)
```

### 3. Literature & Theory (LIT, 15%)

This dimension assesses how you engage with scholarly literature and
the theoretical grounding of your work.

```{r}
#| label: tbl-lit

lit_df <- data.frame(
  Criterion = c(
    "Relevant and sufficient sources",
    "Theoretical framework appropriate",
    "Engagement with literature",
    "Current state of research reflected"
  ),
  Description = c(
    "Are the key works on the topic included, and is coverage sufficient?",
    "Is a coherent theoretical framework chosen and justified?",
    "Is the literature critically engaged with, not merely summarised?",
    "Are recent publications and developments taken into account?"
  )
)

flextable(lit_df) |>
  theme_rubric() |>
  width(j = "Criterion", width = 2.8) |>
  width(j = "Description", width = 3.8)
```

### 4. Results & Analysis (RES, 20%)

This dimension assesses how findings are presented and interpreted.

```{r}
#| label: tbl-res

res_df <- data.frame(
  Criterion = c(
    "Findings clearly presented",
    "Analysis depth and rigor",
    "Results address the research question",
    "Appropriate use of data / material"
  ),
  Description = c(
    "Are the results presented in a clear and accessible way?",
    "Does the analysis go beyond the superficial and show thoroughness?",
    "Do the findings relate clearly to the research question posed at the outset?",
    "Is the data or material used and interpreted appropriately?"
  )
)

flextable(res_df) |>
  theme_rubric() |>
  width(j = "Criterion", width = 2.8) |>
  width(j = "Description", width = 3.8)
```

### 5. Critical Reflection (REF, 10%)

This dimension assesses your ability to evaluate your own work and consider
alternative perspectives.

```{r}
#| label: tbl-ref

ref_df <- data.frame(
  Criterion = c(
    "Self-critical assessment of own work",
    "Awareness of alternative interpretations",
    "Discussion of broader implications",
    "Honest treatment of limitations"
  ),
  Description = c(
    "Are the strengths and weaknesses of your own work openly acknowledged?",
    "Are alternative interpretations of the results considered?",
    "Are wider questions and consequences discussed?",
    "Are limitations dealt with openly rather than concealed?"
  )
)

flextable(ref_df) |>
  theme_rubric() |>
  width(j = "Criterion", width = 2.8) |>
  width(j = "Description", width = 3.8)
```

### 6. Academic Craft (CRF, 15%)

This dimension assesses formal aspects such as language, structure, citation
practice, and formatting.

```{r}
#| label: tbl-crf

crf_df <- data.frame(
  Criterion = c(
    "Structure and coherence (red thread)",
    "Language and expression",
    "Citation and referencing",
    "Formatting, figures, indices"
  ),
  Description = c(
    "Is the work logically structured with a clear thread running through it?",
    "Is the language precise, clear, and appropriate for an academic context?",
    "Are citations consistent and correct, and is the bibliography complete?",
    "Do formatting, figures, and indices meet the stated requirements?"
  )
)

flextable(crf_df) |>
  theme_rubric() |>
  width(j = "Criterion", width = 2.8) |>
  width(j = "Description", width = 3.8)
```

### 7. Independence & Difficulty (IND, 5%)

This dimension recognises intellectual independence and the ambition of the
chosen topic.

```{r}
#| label: tbl-ind

ind_df <- data.frame(
  Criterion = c(
    "Ambition of topic",
    "Originality of approach",
    "Independent intellectual contribution"
  ),
  Description = c(
    "How challenging is the chosen topic?",
    "Is there an independent, creative approach to the subject?",
    "Does the work make its own intellectual contribution beyond mere summary?"
  )
)

flextable(ind_df) |>
  theme_rubric() |>
  width(j = "Criterion", width = 2.8) |>
  width(j = "Description", width = 3.8)
```

## How the Overall Grade Is Calculated

The overall score is the weighted sum of all dimension scores:

$$
\text{Overall score} = \sum_{d=1}^{7} w_d \cdot s_d
$$

where $w_d$ is the weight and $s_d$ the score (0--3) for dimension $d$.
The overall score thus ranges from 0 to 3.
The mapping from scores to letter grades follows the grading key of the
respective course.
